### 2008-12-18 talk/ecoop2009/benchmarks.tex
kill a footnote, add a reference

\begin{itemize}
\item The object is still allocated on the heap, and thus there is an extra
  level of indirection to access the \lstinline{value} field.
\item The method call is optimized through a \emph{polymorphic inline cache}
  (XXX: citation needed), that requires a guard check at each iteration.
\end{itemize}

@@

Despite being only a microbenchark, this result is very important as it proves
that our strategy of intermixing compile time and runtime can yield to better
performances than current techniques.  The result is even more impressive if
we consider dynamically typed languages as TLC are usually considered much
slower than the statically typed ones.

>>>>

\begin{itemize}
\item The object is still allocated on the heap, and thus there is an extra
  level of indirection to access the \lstinline{value} field.
\item The method call is optimized through a \emph{polymorphic inline cache}
  \cite{hoelzle_optimizing_1991}, that requires a guard check at each iteration.
\end{itemize}

@@

Despite being only a microbenchmarks, this result is very important as it proves
that our strategy of intermixing compile time and runtime can yield to better
performances than current techniques.  The result is even more impressive if
we consider dynamically typed languages as TLC are usually considered much
slower than the statically typed ones.

### 2008-12-18 talk/ecoop2009/rainbow.tex
kill a footnote, add a reference

A very different point of view on promotion is as a generalization of techniques
that already exist in dynamic compilers as found in modern virtual machines for
object-oriented language.  In this context feedback
techniques are crucial for good results.  The main goal is to
optimize and reduce the overhead of dynamic dispatching and indirect
invocation.  This is achieved with variations on the technique of
polymorphic inline caches \cite{XXX}: the dynamic lookups are cached and
the corresponding generated machine code contains chains of
compare-and-jump instructions which are modified at run-time.  These
techniques also allow the gathering of information to direct inlining for even
better optimization results.

In the presence of promotion, dispatch optimization can usually be
reframed as a partial evaluation task.  Indeed, if the type of the
object being dispatched to is known at compile-time, the lookup can be
folded, and only a (possibly even inlined) direct call remains in the
generated code.  In the case where the type of the object is not known
at compile-time, it can first be read at run-time out of the object and
promoted to compile-time.  As we will see in the sequel, this produces
machine code very similar to that of polymorphic inline
caches\footnote{Promotion can also be seen as a generalization of
a partial evaluation transformation called "The Trick" (see e.g. \cite{XXX}),
which again produces similar code but which is only applicable for finite sets
of values.}.

>>>>

A very different point of view on promotion is as a generalization of techniques
that already exist in dynamic compilers as found in modern virtual machines for
object-oriented language.  In this context feedback
techniques are crucial for good results.  The main goal is to
optimize and reduce the overhead of dynamic dispatching and indirect
invocation.  This is achieved with variations on the technique of
polymorphic inline caches \cite{hoelzle_optimizing_1991}: the dynamic lookups are cached and
the corresponding generated machine code contains chains of
compare-and-jump instructions which are modified at run-time.  These
techniques also allow the gathering of information to direct inlining for even
better optimization results.

In the presence of promotion, dispatch optimization can usually be
reframed as a partial evaluation task.  Indeed, if the type of the
object being dispatched to is known at compile-time, the lookup can be
folded, and only a (possibly even inlined) direct call remains in the
generated code.  In the case where the type of the object is not known
at compile-time, it can first be read at run-time out of the object and
promoted to compile-time.  As we will see in the sequel, this produces
machine code very similar to that of polymorphic inline
caches.
